### 2012-08-04 docs/dissertation/MSc-coursework.tex
docs

\documentclass[12pt,a4paper]{report}

\usepackage{epcc}
\usepackage{graphics}

% This example file shows how a thesis can be laid out using Latex. It
% does not use any special local features so should be portable to other
% places.
% 
% When producing draft copies of a thesis you may want to print only
% selected pages of the thesis. To do this use the command
% 
% dvips -f -p 4 -n 3 myfile.dvi | lpr
% 
% where -p 4 means start printing at page 4 (ie the page that will be
% numbered 4, not necessarily the 4th page) and -n 3 means print 3 pages.
% This example will print pages 4, 5 and 6.
% 
% If you want to print the thesis and also save paper you can print more
% than one page on each sheet of paper. Use the command
% 
% dvips -f myfile.dvi | psnup -2 | lpr
% 
% to print 2 pages per sheet. psnup can take values 2, 4, 8, or 9.
%
% To produce a PDF version you can create a PostScript copy first
%
% dvips -f myfile.dvi > myfile.pdf
%
% and then convert it
%
% distill myfile.ps
%
% or you can go straight to PDF
%
% pdflatex myfile
%
% Note that pdflatex expects all included figures to be in PDF too. See
% the includegraphics command below.

% This document contains many cross-references and forward references,
% eg in constructing a table of contents, so Latex may need to be run
% twice to get all the references correct. If you need to run Latex twice
% you may get the warning:
% 
% LaTeX Warning: Label(s) may have changed. Rerun to get cross-refSerences right.

% the following 4 lines are the content of the smallmargins.sty file
% but including them explicitly makes this more portable.
%AC%\oddsidemargin=0.1in
%AC%\topmargin=-0.5in
%AC%\textheight=9in
%AC%\textwidth=6.25in

%AC%\parskip 10pt
%AC%\parindent 0in

\begin{document}

%AC%\pagestyle{myheadings}
%AC%\markright{D.~S.~Henty}

%\title{A Latex thesis example}
%\author{D.~S.~Henty}
%\date{\today}

%\maketitle

\pagenumbering{roman}

\title{Hawk-i HPC CLoud Benchmark Tool}
\author{Visakh Muraleedharan}
\date{\today}

\makeEPCCtitle

\thispagestyle{empty}

\vspace{12cm}

\begin{center}

\large{MSc in High Performance Computing}

\large{The University of Edinburgh}

\large{Year of Presentation: 2012}

\end{center}

\newpage

\begin{abstract}
Scientific computing unravel the mysteries of science by constructing mathematical models and numerical algorithms. This required massive computational power and High-Performance Computing (HPC) solutions like cluster and grid have been the answer for massive computing to these needs for a long time. These HPC solutions are hard and expensive to setup, maintain and use. \emph{Cloud computing} is a model of delivering the existing compute infrastructure where computation and storage can be dynamically provisioned on a pay as you go model. Using a vendor cloud service like Amazon Web Service (AWS) can significantly reduce the effort to access these on-demand high performance resources.

In this work we create a platform to study how useful Amazon EC2 cloud computing can be for scientific applications. 
We classify the applications based on the concept of computational \emph{motif}\cite{asanovic2008parallel}  
\end{abstract}

\pagenumbering{roman}

\tableofcontents
\listoftables
\listoffigures

\begin{titlepage}
\vspace*{2in}
% an acknowledgements section is completely optional but if you decide
% not to include it you should still include an empty {titlepage}
% environment as this initialises things like section and page numbering.
\section*{Acknowledgements}

\end{titlepage}

\pagenumbering{arabic}

\chapter{Introduction}

\chapter{Background}

\section{Cloud Computing}
In 1960 John McCarthy predicted that "computation may someday be organised as a public utility". Availability of high  capacity networks and low cost hardware as well as the adoption of hardware virtualisation technology made this prediction come true. This delivery model enabled a large community of end-users to obtain computing and storage capacity on-demand.

\subsection{Why Cloud Computing?}
\subsubsection{Resource on-demand}
easy, accesibility
\subsubsection{Pay as you go}
\subsubsection{Maintenance and Upgradation}
\subsubsection{Eco-friendly}

\subsection{Cloud Computing service models}
\subsubsection{IaaS}
\subsubsection{PaaS}
\subsubsection{Saas}

\subsection{Cloud Computing Deployment models}
\subsubsection{Private Cloud}
\subsubsection{Public Cloud}
\subsubsection{Hybrid Cloud}
\subsubsection{Community Cloud}

\section{Amazon Web Service}
\subsection{Amazon Machine Instance}
\subsection{Types of Instances}
\subsection{Amazon EC2}
different instances chart and limitations of microinstance.
\subsubsection{EC2 cluster compute}

\section{HPC in the cloud}
Introduction Paragraph, comparison with Clusters
\subsection{Advantages}
\subsection{Limitation}
Network Limitations, Availability, Stability of performance
Benifits
\subsection{Amazon Cluster compute}

\section{Parallel Applications and 13 Dwarfs}
Explain the 13 dwarfs and under the two sections explain N-body and FFT

\subsection{N-Body Methods}
MD, About the dwarf and application types optimisation, n2 and nlogn
Communication pattern!!!

\subsection{Spectral Methods}
In scientific computing, a class of techniques used to numerically solve differential equations involving the use of Fast Fouries Transform are called spectral methods

Fourier transform
Jean Baptiste Joseph Fourier (1768-1830) first employed what we now call Fourier transforms whilst working on the theory of heat 
Linear transform which takes temporal or spatial information and converts into information which lies in the frequency domain
Who would use Fourier Transforms? 
• Physics 
– Cosmology (P3M N-body solvers) 
– Fluid mechanics 
– Quantum physics 
– Signal and image processing 
– Antenna studies 
– Optics

• Numerical analysis 
– Linear systems analysis 
– Boundary value problems 
– Large integer multiplication (Prime finding)

• Statistics 
– Random process modelling 
– Probability theory

{\underline Discrete Fourier Transform}
The discrete Fourier transform of N complex points fk is defined as

Communication pattern!!!

\chapter{Live Benchmark Tool Setup}
\section{System Design}
\subsection{Architecture}
\subsection{Sequence diagram}
\subsection{Database Design}

\section{Instance types Used}
Refer to graph in Background
Why these types?
\section{Sun Grid Engine Cluster}
Why sungrid Engine
\section{SGE Clustering in Amazon cloud}
  benchmark to show how cluster instances are faster than normal high cpu instances
\section{Web interface}
Describe in detail
\subsection{Admin panel}
\subsection{User Dashboard}

\chapter{Results and Analysis}
\section{Serial}
for each instance type
time to result, execution time, increasing problem size
N-body and FFT

\section{Parallel program}
execution time, increasing problem size, number of cores
N-body and FFT
\section{Stability of results}
History of execution, snapshot from dashboard

\chapter{Conclusions}

\appendix
% the appendix command just changes heading styles for appendices.

\chapter{Cluster Computing Setup}
\subsection{Using Starcluster}
\subsection{Cluster Management package}
\subsection{Using Sun Grid Engine}
\subsection{Creating Dashboard}
\chapter{Benchmarking  programs}
\subsection{Spectral Methods}
\subsection{N-body}

\bibliographystyle{plain}
\bibliography{dissertation}

\end{document}